Theoretical prediction of disrupted Min oscillation in flattened \emph{Escherichia coli}

### 2015-07-01 paper/response-to-referees.txt

1. The largest concern of Referee #1 was our choice of Huang et al's
model for our simulations.  The referee acknowledges that we can stay
with this model, but feels that our justifaction that this model is
inadequate.  The referee objects that the fact that this model has
been widely is an inappropriate justification.  While we agree that
this alone is not a sufficient reason for studying this model, we do
believe that it is relevant: this is not some obscure older model that
we are testing, but rather the basis of all more recent models.  We
believe that the limitations of this model are of interest, and have
added some discussion to the end of the introduction section
explaining this.

2. As a secondary point, the referee points out we cannot conclude
that some other deterministic model could not demonstrate chaotic
behavior that explains the experimental results of Mannik et al.  We
agree that this cannot be ruled out, and have added mention of this to
our "Comparison of Experimental and Stadium Shapes", and have made
even more tentative our suggestion in the summary that other similar
deterministic models may also prove to be non-chaotic.

3. The referee when looking at Figure 4 observes that the stadium
shapes exhibit osciallation that is largely bipolar, and expresses
doubt in our claim that the oscillations in stadium shapes are
irregular.  In particular, the referee points out that if we average
over many periods, we would find that the oscillations are indeed
bipolar.  We agree with this statement, but not with the conclusion.
Averaging the locations of the maxima would indeed miss the point of
the comparison with experiment, which is in examining the sequence of
locations of maxima.

We have edited the "comparison of experimental and stadium shapes"
section, the conclusion, and the abstract to moderate our claims regarding the spatial irregularity of the
experimental and stadium shapes as predicted by the stochastic model.

@@

5. Based on the referee's suggestion, we have clarified that reference
11 does not extend Huang's model, except in that it simulates the same
reaction-diffusion equations stochastically.

6. The referee points out that we failed to mention that stochastic
simulations are needed to explain the experimentally observed
length-dependence of fluctuations in the Min system.  We have added
mention of this additional evidence of the importance of inclusion of
stochastic effects, in the paragraph suggested by the referee.

Referee #3.

7. The referee #3 also expressed doubt about our conclusion with
regard to the effect of irregular shape which we base on our stadium
models.  Unlike the previous referee, the criticism is based not on
the spatial output (Figure 4), but rather on the reported coherence
times.  We agree with the referee that our concluding sentence
immediately prior to the conclusion section was unjustified, and have
edited this sentence to reflect the referee's objections.
JEFFSAYS{The only thing I could see to change in addition to the above is that 
I added a 'maybe' to the abstract.  Don't think we need to note it.}

@@

<Jeff will check if we talk about decoherence times later>

8. The referee also questioned how the cell was divided in order to
perform the coherence time calculations, pointing out that a smaller
region could lead to a shorter coherence time.  EXPLAIN HOW WE DIVIDE
THE CELLS UP IN THIRDS BY VOLUME? IT'S AWKWARD THAT WE DON'T DIVIDE
ALONG LONG AXIS.  JEFFSAYS{Bad news here.  It looks like I did this
dividing up the cells by eye.  Originally the slices were used to
created the box plots, which was only a way to visualize the data in
smaller time domains, so there's wasn't really any disadvantage to
doing it roughly like this (we even had the visually appealing little
colored diagrams on the sides of the box plots).  Not sure what to say
abou this in text.}
JEFFSAYS {I figured we should do this one together because I really didn't
know what to say.  Here's the volumes:
shape A - left 1.29 mid 3.29 right 2.12
shape B - left 1.41 mid 2.48 right 1.91
stad A - left 1.91 mid 2.62 right 1.84
stad B - left 1.88 mid 2.49 right 1.82}

### 2015-07-01 paper/plos.tex
more changes, added volumes, addressed stadium irregularity problem, and added a maybe about flattening in abstract

* E-mail: Corresponding schuljef@physics.oregonstate.edu
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
  The dynamics of the Min-protein system help \emph{Escherichia coli}
  regulate the process of cell division by identifying the center of
  the cell.  While this system exhibits robust bipolar oscillations in
  wild-type cell shapes, recent experiments have shown that when the
  cells are mechanically deformed into wide, flattened out, irregular
  shapes, the spatial regularity of these oscillations breaks down. We
  employ widely used stochastic and deterministic models of the Min
  system to simulate cells with flattened shapes.  The deterministic
  model predicts strong bipolar oscillations, in contradiction with
  the experimentally observed behavior, while the stochastic model,
  which is based on the same reaction-diffusion equations, does
  predict spatially irregular oscillations as observed in experiment\fixme{, although they are generally more bipolar}.
  We further report simulations of flattened but more symmetric
  shapes, and find that it \fixme{may be?} is the flattening and lateral expansion
  rather than the asymmetry of the cell shapes that causes the
  irregular oscillation behavior.

@@

\section*{Conclusion}
\fixme{We find a certain amount of qualitive agreement between our stochastic simulations and experiments showing disrupted MinD oscillations in asymmetric flattened E. coli cells.}We find agreement between our stochastic simulations and experiments
showing disrupted bipolar MinD oscillation in asymmetric flattened
E. coli cells~\cite{mannik2012robustness}.  As observed
experimentally, our simulations predict that MinD maxima will form in
a spatially irregular sequence of locations, \fixme{albeit with a with more course scale bipolar regularity.}  This result builds upon
existing results showing that this model and its variants are
effective in a variety of wild-type and mutated cell
shapes~\cite{fange2006noise, varma2008min, kruse2007experimentalist},
and reinforces that stochastic variations of Huang's 2003
model~\cite{fange2006noise, kerr2006division} has considerable
predictive power beyond the standard wild-type cell.

>>>>

* E-mail: Corresponding schuljef@physics.oregonstate.edu
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
  The dynamics of the Min-protein system help \emph{Escherichia coli}
  regulate the process of cell division by identifying the center of
  the cell.  While this system exhibits robust bipolar oscillations in
  wild-type cell shapes, recent experiments have shown that when the
  cells are mechanically deformed into wide, flattened out, irregular
  shapes, the spatial regularity of these oscillations breaks down. We
  employ widely used stochastic and deterministic models of the Min
  system to simulate cells with flattened shapes.  The deterministic
  model predicts strong bipolar oscillations, in contradiction with
  the experimentally observed behavior, while the stochastic model,
  which is based on the same reaction-diffusion equations, does
  predict spatially irregular oscillations\fixme{, although they are still somewhat more bipolar than seen in experiment}.
  We further report simulations of flattened but more symmetric
  shapes, and find that it \fixme{may be} the flattening and lateral expansion
  rather than the asymmetry of the cell shapes that causes the
  irregular oscillation behavior.

@@

\section*{Conclusion}
\fixme{We find a certain amount of qualitive agreement between our stochastic simulations and experiments showing disrupted MinD oscillations in asymmetric flattened E. coli cells.}  We find agreement between our stochastic simulations and experiments
showing disrupted bipolar MinD oscillation in asymmetric flattened
E. coli cells~\cite{mannik2012robustness}.  As observed
experimentally, our simulations predict that MinD maxima will form in
a spatially irregular sequence of locations, \fixme{albeit with a somewhat more course scale bipolar regularity.}  This result builds upon
existing results showing that this model and its variants are
effective in a variety of wild-type and mutated cell
shapes~\cite{fange2006noise, varma2008min, kruse2007experimentalist},
and reinforces that stochastic variations of Huang's 2003
model~\cite{fange2006noise, kerr2006division} has considerable
predictive power beyond the standard wild-type cell.
